---
title: "Plaque_Species_level_RA_clr splsda"
output: html_document
date: '2022-09-14'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown



## R Markdown

```{r}
rm(list = ls())
```

# libraries General

```{r}
library(tidyverse)
library(magrittr)
library(DT)
library(phyloseq)
library(ggplotify)
library(cowplot)
```

# Libraries Speciesecific

```{r}
library(PCAtools) # for the command PCAtools::pca
library(mixOmics)
library(phyloseq)
library(microbiome)
library(PLSDAbatch)
library(vegan) #varpart
```

```{r}
in_dir  = "/home/wasif/Rproj_221101_Plaque_Genus_Species_metaanalysis/Plaque_Species/Species_221101_2_processed_data_files"
out_dir = "/home/wasif/Rproj_221101_Plaque_Genus_Species_metaanalysis/Plaque_Species/Species_221101_2_processed_data_files"
```

# Import data

```{r}
Pooled_Pl_Species_RA_table = read.table(paste0(in_dir, "/Plaque_Species_relative_abundance_OTU_table_with_meta_data.txt"),
                                   header = T,
                                   sep = "\t",
                                  check.names = F )
Pooled_Pl_Species_RA_table
```

# separate OTU and metadata

```{r}
# Meta
Pooled_Pl_Species_meta = Pooled_Pl_Species_RA_table %>% 
  column_to_rownames("sampleid") %>% 
  dplyr::select(c(Disease_status ,Host_age, Host_sex, Sample_type, Geo_location, Study_name))
Pooled_Pl_Species_meta

# OTU
# Pooled_Pl_Species_RA_t = Pooled_Pl_Species_RA_table %>% 
#   column_to_rownames("sampleid") %>% 
#   dplyr::select(!c(Disease_status ,Host_age, Host_sex, Sample_type, Geo_location, Study_name)) %>% 
#   t() %>% 
#   as.data.frame()
# Pooled_Pl_Species_RA_t
```

# import clr data
```{r}
load(paste0(out_dir, "/Pooled_Pl_Species_RA_clr.RData"))
Pooled_Pl_Species_RA_clr
```

#  pca option

```{r}

Pooled_Pl_Species_RA_OTU_mixOmics_pca_clr <- mixOmics::pca(Pooled_Pl_Species_RA_clr, scale = T,
                                                            center = T,
                                                            ncomp = 3,
                                                            )
```

```{r}
plot_before_splsda = 
mixOmics::plotIndiv(Pooled_Pl_Species_RA_OTU_mixOmics_pca_clr,
                    group=Pooled_Pl_Species_meta$Study_name, 
                    ind.names = F,
                    legend=T,
                    point.lwd = 0.7,
                    pch = as.numeric(factor(Pooled_Pl_Species_meta$Disease_status))+14,
                    pch.levels=(Pooled_Pl_Species_meta$Disease_status), ellipse = TRUE, 
                    title="Before batch correction",legend.title = "Study_name", 
                    legend.title.pch = "Disease_status",
                    size.legend = rel(2.4),
                    size.legend.title = rel(1.0),
                    size.title = rel(1.2)
                    )
```
```{r}
plot_before_splsda_gg = plot_before_splsda$graph +theme(legend.position="none")
```


# check if the row names in metadata and OTU files are the same

```{r}
identical(row.names(Pooled_Pl_Species_meta), row.names(Pooled_Pl_Species_RA_clr))
```

# Set batches and disease_status

```{r}
Pooled_Disease_status = as.factor(Pooled_Pl_Species_meta$Disease_status)
Pooled_Disease_status
Pooled_Study_name = as.factor(Pooled_Pl_Species_meta$Study_name)
Pooled_Study_name
```


# The optimal number of components

```{r}
Pooled_Pl_Species_RA_clr_Disease_status.tune <- mixOmics::plsda(X = Pooled_Pl_Species_RA_clr, Y = Pooled_Disease_status, ncomp = 5)
Pooled_Pl_Species_RA_clr_Disease_status.tune$prop_expl_var #1
```

```{r}
Pooled_Pl_Species_RA_clr_Study_name.tune <- PLSDA_batch(X = Pooled_Pl_Species_RA_clr, 
                             Y.trt = Pooled_Disease_status, Y.bat = Pooled_Study_name,
                             ncomp.trt = 1, ncomp.bat = 10)
Pooled_Pl_Species_RA_clr_Study_name.tune$explained_variance.bat  #3 # the sum to top 2 Y values give a variance of 1 #see the next code 
```

```{r}
sum(Pooled_Pl_Species_RA_clr_Study_name.tune$explained_variance.bat$Y[1:4])
```

# To correct the Pooled_Pl_Species_RA_clr\_ data with PLSDA-batch

```{r}
Pooled_Pl_Species_RA_clr_plsda_batch.correct <- PLSDA_batch(X = Pooled_Pl_Species_RA_clr, 
                                      Y.trt = Pooled_Disease_status, Y.bat = Pooled_Study_name, 
                                      ncomp.trt = 1, ncomp.bat = 4)
Pooled_Pl_Species_RA_clr_plsda_batch <- Pooled_Pl_Species_RA_clr_plsda_batch.correct$X.nobatch
Pooled_Pl_Species_RA_clr_plsda_batch %>% datatable()
```

# sPLSDA-batch The optimal number of variables to select per component

```{r}
set.seed(789)
Pooled_Pl_Species_RA_clr_test.keepX = c(seq(1, 10, 1), seq(20, ncol(Pooled_Pl_Species_RA_clr), 10)) #number of OTUs
Pooled_Pl_Species_RA_clr_test.keepX
Pooled_Pl_Species_RA_clr_Disease_status.tune.v <- tune.splsda(X = Pooled_Pl_Species_RA_clr, Y = Pooled_Disease_status, 
                             ncomp = 1, test.keepX = Pooled_Pl_Species_RA_clr_test.keepX, 
                             validation = 'Mfold', folds = 4, nrepeat = 50)
Pooled_Pl_Species_splsda_choice_keepX = Pooled_Pl_Species_RA_clr_Disease_status.tune.v$choice.keepX 
Pooled_Pl_Species_splsda_choice_keepX
```

```{r}
Pooled_Pl_Species_RA_clr_splsda.correct <- PLSDA_batch(X = Pooled_Pl_Species_RA_clr, 
                                       Y.trt = Pooled_Disease_status, Y.bat = Pooled_Study_name, 
                                       ncomp.trt = 1, keepX.trt = Pooled_Pl_Species_splsda_choice_keepX, 
                                       ncomp.bat = 4)
Pooled_Pl_Species_RA_clr_splsda <- Pooled_Pl_Species_RA_clr_splsda.correct$X.nobatch
Pooled_Pl_Species_RA_clr_splsda %>%  datatable()
```
# Save the sPLSDA batch corrected table
```{r}
save(Pooled_Pl_Species_RA_clr_splsda, file = paste0(out_dir, "/Pooled_Pl_Species_RA_clr_splsda.RData"))
```


#  pca option

```{r}

Pooled_Pl_Species_RA_clr_splsda_pca <- mixOmics::pca(Pooled_Pl_Species_RA_clr_splsda, scale = T,
                                                            center = T,
                                                            ncomp = 3,
                                                            )
```

```{r}
plot_after_splsda = 
mixOmics::plotIndiv(Pooled_Pl_Species_RA_clr_splsda_pca,
                    group=Pooled_Pl_Species_meta$Study_name, 
                    ind.names = F,
                    legend=T,
                    point.lwd = 0.7,
                    pch = as.numeric(factor(Pooled_Pl_Species_meta$Disease_status))+14,
                    pch.levels=(Pooled_Pl_Species_meta$Disease_status),
                    ellipse = TRUE, 
                    title="After sPLSDA batch correction",
                    legend.title = "Study_name", 
                    legend.title.pch = "Disease_status",
                    size.legend = rel(2.4),
                    size.legend.title = rel(1.0),
                    size.title = rel(1.2),
                    legend.position = "bottom",
                    )
```
```{r}
plot_after_splsda
```

```{r}
plot_after_splsda_gg = plot_after_splsda$graph +theme(legend.position="none")
```

```{r}

legend = get_legend( plot_after_splsda$graph +theme(legend.position="top"))

# build grid without legends
pgrid <- plot_grid(plot_before_splsda_gg, plot_after_splsda_gg, ncol = 2)
# add legend
p <- plot_grid(
  legend,
  plot_grid(plot_before_splsda_gg, plot_after_splsda_gg, ncol = 2),
  ncol = 1,
  rel_heights = c(0.15, 1) # Adjust these values to change the relative heights of the legend and plots
)
p
ggsave("before_n_after_splsda.png",
       dpi = 300,
       height = 5,
       width = 10)
```
```{r}
legend
```

































