---
title: "Species level modified microbiomeR ggplot for metanalysis on sPLSDA data"
output: html_document
date: '2022-09-20'
---

# Additional comments related to metamicrobiomeR

1.  metamicrobiomeR does not work with the original Species name with brackets eg. Absconditabacteria\_[SR1]\_(G.1)\_bacterium_HMT_345. This can be taken care by 'check.names = T' option.
2.  metamicrobiome need prefix k in front of taxonomic names

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
rm(list = ls())
```

# Libraries

```{r}
library(tidyverse)
library(DT)
library(plyr)
library(gdata)
library(ggpubr)
library(cowplot)
library(egg)
library(phyloseq)
```

# Libraries

```{r}
library(metamicrobiomeR)
library(phyloseq)
library(microbiome)
```

# Import data

```{r}
in_dir  = "../../Plaque_Species/Species_221101_2_processed_data_files"
#out_dir = "/home/wasif/Rproj_221101_Plaque_Species_Species_metaanalysis/Plaque_Species/Species_221101_2_processed_data_files"
```

# Import data

```{r}
Pooled_Pl_Species_RA_table = read.table(paste0(in_dir, "/Plaque_Species_relative_abundance_OTU_table_with_meta_data.txt"),
                                   header = T,
                                   sep = "\t",
                                  check.names = F )
Pooled_Pl_Species_RA_table
```

# Load the RA_clr_splsda file

```{r}
load(paste0(in_dir, "/Pooled_Pl_Species_RA_clr_splsda.RData"))
Pooled_Pl_Species_RA_clr_splsda

Pooled_Pl_Species_RA_clr_splsda = Pooled_Pl_Species_RA_clr_splsda %>% as.data.frame()
Pooled_Pl_Species_RA_clr_splsda

# modify OTU names according to metamicrobiomeR
Pooled_Pl_Species_RA_clr_splsda = Pooled_Pl_Species_RA_clr_splsda %>% 
  setNames(paste0('k__s_', names(.)))
Pooled_Pl_Species_RA_clr_splsda
```

# separate OTU and metadata

```{r}
# Meta
Pooled_Pl_Species_RA_table_meta = Pooled_Pl_Species_RA_table %>% 
  column_to_rownames("sampleid") %>% 
  dplyr::select(c(Disease_status ,Host_age, Host_sex, Sample_type, Geo_location, Study_name))
Pooled_Pl_Species_RA_table_meta

# # OTU
# Pooled_Pl_Species_RA_table_OTU = Pooled_Pl_Species_RA_table %>% 
#   column_to_rownames("sampleid") %>% 
#   dplyr::select(!c(Disease_status ,Host_age, Host_sex, Sample_type, Geo_location, Study_name)) %>% 
#   setNames(paste0('k__s_', names(.))) # metamicrobiome need prefix k in front of taxonomic names
# Pooled_Pl_Species_RA_table_OTU
```

```{r}
Pooled_Pl_Species_RA_clr_splsda_meta_OTU = merge(Pooled_Pl_Species_RA_table_meta[ , c(1,6)],
                                          Pooled_Pl_Species_RA_clr_splsda, by = "row.names") %>%
  column_to_rownames("Row.names") 
Pooled_Pl_Species_RA_clr_splsda_meta_OTU
```

# apply metamicrobiomeR::taxa.compare with for loop on each study

```{r include=FALSE}
Studies =  unique(Pooled_Pl_Species_RA_clr_splsda_meta_OTU$Study_name) %>% sort()
Studies
Taxa_compare_res = list()
for (Study in Studies) {
  Pooled_Pl_Species_RA_clr_splsda_meta_study = Pooled_Pl_Species_RA_clr_splsda_meta_OTU %>% 
    filter(Study_name == Study) %>% 
    dplyr::select(!Study_name)
  # apply metamicrobiomeR::taxa.compare
  Taxa_compare_res_study = taxa.compare(taxtab=Pooled_Pl_Species_RA_clr_splsda_meta_study,
                         propmed.rel="lm",
                         comvar="Disease_status",
                         adjustvar=NULL,
                         longitudinal="no",
                         p.adjust.method="fdr")
  Taxa_compare_res_study_2 = Taxa_compare_res_study %>% 
    mutate(study =  Study) %>% 
    mutate(pop = Study)
  Taxa_compare_res[[Study]] = Taxa_compare_res_study_2
}
```

# find common rows after taxa compare. to know the Species present in all joined datasets

<https://www.r-bloggers.com/2012/06/intersect-for-multiple-vectors-in-r/> <https://stackoverflow.com/questions/13016359/how-to-directly-select-the-same-column-from-all-nested-lists-within-a-list>

```{r}
# For Inner Join
Species_intersection <- Reduce( union, lapply(Taxa_compare_res, `[`, 1) )
Species_intersection
```

```{r}
Pooled_Pl_Species_RA_clr_Taxa_compare_res = bind_rows(Taxa_compare_res) %>% 
  #filter(id %in% "k__bacteria.p__P.c__C.o__O.f__F.g__Actinomyces_dentalis")
  filter(id %in% Species_intersection$id)
Pooled_Pl_Species_RA_clr_Taxa_compare_res
```

# meta analysis

```{r}
Pooled_Pl_Species_RA_clr_meta_taxa= meta.taxa(taxcomdat=Pooled_Pl_Species_RA_clr_Taxa_compare_res,
                                              summary.measure="RD",
                                              pool.var="id",
                                              studylab="study",
                                              backtransform =FALSE,
                                              percent.meta=0.4,
                                              p.adjust.method="fdr")
Pooled_Pl_Species_RA_clr_meta_taxa$random$Disease_statusECC
```

```{r}
# metatab.show(metatab = Pooled_Pl_Species_RA_clr_meta_taxa$RA_clrndom,
#                       com.pooled.tab = Pooled_Pl_Species_RA_clr_Taxa_compare_res,
#                       #highest.lev = "p",
#              sumvar = "path",
#                       #tax.lev="l5",
#                       showvar="Disease_statusECC",
#                       p.cutoff.type="p",
#                       p.cutoff=1,
#                       display="table"
#                       )

Pooled_Pl_Species_RA_clr_metadat<-metatab.show(metatab = Pooled_Pl_Species_RA_clr_meta_taxa$random,
                      com.pooled.tab = Pooled_Pl_Species_RA_clr_Taxa_compare_res,
                      highest.lev = "s",
             sumvar = "path",
                      tax.lev="l1",
                      showvar="Disease_statusECC",
                      p.cutoff.type="p.adjust",
                      p.cutoff=0.04,
                      display="data"
                      )
Pooled_Pl_Species_RA_clr_metadat
```

```{r}

meta.niceplot(metadat=Pooled_Pl_Species_RA_clr_metadat,
              sumtype="taxa",
              level="main",
              p="p",
              p.adjust="p.adjust",
              phyla.col="black",
              p.sig.heat="yes",
              heat.forest.width.ratio =c(1,1.5),
              leg.key.size=0.6,
              leg.text.size=8,
              heat.text.x.size=8,
              heat.text.x.angle=10,
              forest.axis.text.y=10,
              forest.axis.text.x=8,
              point.ratio = c(4,2),
              line.ratio = c(2,1)
              )

```

```{r}
Pooled_Pl_Species_RA_clr_metadat
```

```{r}
source("../../R_functions_wk/meta_niceplot_wk.R")

png("./Pooled_Pl_Species_RA_clr_meta_niceplot.png",
    width = 6,
       height = 3,
    units = "in",
    res = 600)
meta_niceplot_wk(metadat=Pooled_Pl_Species_RA_clr_metadat,
              sumtype="taxa",
              level="main",
              p="p",
              p.adjust="p.adjust",
              phyla.col="black",
              p.sig.heat="yes",
              heat.text.x.angle=45

              )

# ggsave("./Pooled_Pl_Species_RA_clr_meta_niceplot",
#       plot =  Pooled_Pl_Species_RA_clr_meta_niceplot,
#        width = 600,
#        height = 300,
#        units = "px"
# )

dev.off()
```
